\documentclass[../notes.tex]{subfiles}

\pagestyle{main}
\renewcommand{\chaptermark}[1]{\markboth{\chaptername\ \thechapter\ (#1)}{}}

\begin{document}




\chapter{From Classical to Quantum Mechanics}
\section{Blackbody Radiation}
\begin{itemize}
    \item \marginnote{9/27:}The surface of a hot body emits energy in the form of EM radiation.
    \item Changes that occur with temperature:
    \begin{itemize}
        \item If less than $\SI{500}{\celsius}$, we have IR Radiation (heat).
        \item If $\SIrange{500}{600}{\celsius}$, we have visible radiation (a glowing body).
        \item If $\SI{5000}{\celsius}$, we have a "white hot" body (short wavelength).
    \end{itemize}
    \item As a body gets hotter, it emits shorter wavelength radiation.
    \item \textbf{Stefan-Boltzmann law}: The the total emissive power $R$ (recall that power is en / time) of a blackbody (BB) is given by
    \begin{equation*}
        R(T) = \sigma T^4
    \end{equation*}
    where $\sigma\approx\SI{5.67e-8}{\watt\per\square\meter\per\kelvin\tothe{4}}$ is \textbf{Stefan's constant}.
    \begin{itemize}
        \item Work done by Stefan and Boltzmann (c. 1870 / 1884, respectively).
    \end{itemize}
    \item \textbf{Wien's 1st Law}: The wavelength for maximum emissive power obeys the equation
    \begin{equation*}
        \lambda_\text{max}T = b
    \end{equation*}
    where $b=\SI{2.808e-3}{\meter\kelvin}$ is \textbf{Wein's displacement constant}.
    \begin{figure}[h!]
        \centering
        \begin{tikzpicture}
            \footnotesize
            \draw [<->] (0,4.5) -- node[left]{$R(\lambda,T)$} (0,0) -- node[below=2mm]{$\lambda\ (\si{\micro\meter})$} (7.5,0);
            \foreach \x in {1,...,7} {
                \draw (\x,0.1) -- ++(0,-0.2);
            }
            \foreach \y in {1,...,4} {
                \draw (0.1,\y) -- ++(-0.2,0);
            }
            
            \begin{scope}[xscale=1.8]
                \draw [blx,thick,samples=100,smooth,/pgf/fpu,/pgf/fpu/output format=fixed] plot[domain=0.1:4] (\x,{5*10^2/(\x^5*(e^(9900/(2000*\x))-1))});
                \draw [grx,thick,samples=100,smooth,/pgf/fpu,/pgf/fpu/output format=fixed] plot[domain=0.1:4] (\x,{5*10^2/(\x^5*(e^(9900/(1750*\x))-1))});
                \draw [orx,thick,samples=100,smooth,/pgf/fpu,/pgf/fpu/output format=fixed] plot[domain=0.1:4] (\x,{5*10^2/(\x^5*(e^(9900/(1500*\x))-1))});
                \draw [rex,thick,samples=100,smooth,/pgf/fpu,/pgf/fpu/output format=fixed] plot[domain=0.1:4] (\x,{5*10^2/(\x^5*(e^(9900/(1000*\x))-1))});
        
                \draw [semithick] (0.997,3.667) node[above right]{$\lambda_\text{max}$} -- ++(0,-0.2);
                \draw [semithick] (1.139,1.93) -- ++(0,-0.2);
                \draw [semithick] (1.329,0.946) -- ++(0,-0.2);
                \draw [semithick] (1.994,0.211) -- ++(0,-0.2);
            \end{scope}
    
            \begin{scope}[xshift=6cm,yshift=3cm]
                \node at (0,1.2) {${\color{blx}\blacksquare}=\SI{2000}{\kelvin}$};
                \node at (0,0.8) {${\color{grx}\blacksquare}=\SI{1750}{\kelvin}$};
                \node at (0,0.4) {${\color{orx}\blacksquare}=\SI{1500}{\kelvin}$};
                \node at (0,0) {${\color{rex}\blacksquare}=\SI{1000}{\kelvin}$};
            \end{scope}
        \end{tikzpicture}
        \caption{Wein's 1st Law.}
        \label{fig:WeinLaw1}
    \end{figure}
    \item Area under the curve (found with integration) is the total emissive power.
    \item We now change variables from emissive power $R$ to energy density $\rho$ in the BB cavity.
    \begin{equation*}
        \rho(\lambda,T) = \frac{4}{c}R(\lambda,T)
    \end{equation*}
    \item Wien's 2nd Law (1893): The energy density must have a functional relationship with the following form.
    \begin{equation*}
        \rho(\lambda,T) = \frac{f(\lambda T)}{\lambda^5}
    \end{equation*}
    \begin{itemize}
        \item $f(\lambda T)$ cannot be determined from thermodynamics. Thus, something else is needed!
    \end{itemize}
    \item Lord Rayleigh and his graduate student Jeans (1899) propose a solution.
    \begin{itemize}
        \item EM: The thermal radiation within a cavity must exist in the form of standing waves.
        \item RJ showed that the number $n$ of standing waves per unit volume, per wavelength has the following form.
        \begin{equation*}
            n(\lambda) = \frac{8\pi}{\lambda^4}
        \end{equation*}
        \item If $\bar{\epsilon}$ is the average energy in the mode with wavelength $\lambda$, then
        \begin{equation*}
            \rho(\lambda,T) = \frac{8\pi}{\lambda^4}\bar{\epsilon}
        \end{equation*}
        \item Waves come from atoms in the walls of the BB cavity, which act as linear harmonic oscillators at a frequency $\nu=c/\lambda$.
        \item Assuming thermal equilibrium, we obtain
        \begin{align*}
            \bar{\epsilon} &= \frac{\int_0^\infty\epsilon\e[-\epsilon/kT]}{\int_0^\infty\e[-\epsilon/kT]}\\
            &= -\pdv{\beta}\ln\left( \int_0^\infty\e[-\beta\epsilon]\dd{\epsilon} \right)\\
            &= \frac{1}{\beta}\\
            &= kT
        \end{align*}
        where $k$ is the Boltzmann constant.
        \begin{itemize}
            \item Basically, we sum all energies $\epsilon$, weighted by the probability $\e[-\epsilon/kT]$ of the energy existing, and divided by the total energy.
            \item The first equation is equivalent to the second with $\beta=1/kT$.
        \end{itemize}
        \item Therefore,
        \begin{equation*}
            \rho(\lambda,T) = \frac{8\pi kT}{\lambda^4}
        \end{equation*}
    \end{itemize}
    \item UV catastrophe: Rayleigh's formula diverges from the experimental data for short wavelength.
    \begin{itemize}
        \item The above formula diverges to $+\infty$, driven by the $\lambda^4$ term in the denominator, as $\lambda\to 0$. However, the amount of radiation of shorter wavelengths should decrease past a point, as seen in Figure \ref{fig:WeinLaw1}.
    \end{itemize}
    \item Max Planck comes in, proposes an idea to the German acacdemy that's so radical, they think he's insane, but he's actually right and it lays a key idea for quantum mechanics.
    \item Planck's key insight: The energy levels of the oscillators are not continuous, but are quantized.
    \begin{itemize}
        \item So we can't actually take an integral as Rayleigh did; we have to take an infinite series.
        \item In reality,
        \begin{align*}
            \bar{\epsilon} &= \frac{\sum_{n=0}^\infty n\epsilon_0\e[-\beta n\epsilon_0]}{\sum_{n=0}^\infty\e[-\beta n\epsilon_0]}\\
            &= \frac{\epsilon_0}{\e[\beta\epsilon_0]-1}
        \end{align*}
        \item Thus,
        \begin{equation*}
            \rho(\lambda,T) = \frac{8\pi\epsilon_0}{\lambda^4(\e[\epsilon/kT]-1)}
        \end{equation*}
        \item But to satisfy Wien's 2nd law, we must let $\epsilon_0\propto 1/\lambda$. More specifically, $\epsilon_0=hc/\lambda=h\nu$, where $h$ is Planck's constant.
        \begin{itemize}
            \item This setup allowed us to get an accurate value for Planck's constant for the first time in history.
        \end{itemize}
        \item Planck's theory predicts the data of Figure 1.
    \end{itemize}
    \item A perfect blackbody absorbs and emits radiation at all frequencies.
    \begin{itemize}
        \item A star is pretty close to a blackbody. The graphite in a pencil is 97\% a blackbody. We are all blackbodies.
        \item The entire universe can be viewed as a blackbody.
    \end{itemize}
    \item Princeton and Bell Labs telescopes find \textbf{Cosmic Background Radiation} (A. A. Penzias and R. W. Wilson, 1964).
    \begin{itemize}
        \item Background radiation from the universe itself.
        \item $\lambda_\text{max}=\SI{7.35}{\centi\meter}$.
        \item Isotropic radio signal, that comes form everywhere.
        \item From this, you can workout the temperature of the universe from Wein's first law.
        \item Thus, the whole universe is a blackbody with a temperature of approximately $\SI{3}{\kelvin}$.
    \end{itemize}
\end{itemize}




\end{document}