\documentclass[../notes.tex]{subfiles}

\pagestyle{main}
\renewcommand{\chaptermark}[1]{\markboth{\chaptername\ \thechapter\ (#1)}{}}
\setcounter{chapter}{3}

\begin{document}




\chapter{The Hydrogen Atom and Angular Momentum}
\section{Rotational Motion}
\begin{itemize}
    \item \marginnote{10/18:}Consider the diatomic molecule \ce{AB} at a distance $r_0$ apart rotating about its center of mass.
    \begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.3\linewidth}
            \centering
            \begin{tikzpicture}
                \footnotesize
                \draw (0,0) node[circle,fill=rex,inner sep=1.5pt,label={below:$A$}]{}
                    -- node[pos=0.65,fill,inner sep=1.5pt,label={below:CM}]{}
                    (10:2) node[circle,fill=blx,inner sep=1.5pt,label={below:$B$}]{}
                ;
                \draw [decorate,decoration=brace] (100:0.2) -- node[above=2pt]{$r_0$} ++(10:2);
            \end{tikzpicture}
            \caption{Diatomic.}
            \label{fig:partRota}
        \end{subfigure}
        \begin{subfigure}[b]{0.3\linewidth}
            \centering
            \begin{tikzpicture}
                \footnotesize
                \draw [semithick,-latex] (10:2) -- ++(100:0.5);
                \draw (0,0) node[fill,inner sep=1.5pt,label={left:origin}]{}
                    --
                    (10:2) node[circle,fill=blx!65!rex,inner sep=1.5pt,label={right:$\mu$}]{}
                ;
                \draw [decorate,decoration={brace,mirror}] (-80:0.2) -- node[below=2pt]{$r_0$} ++(10:2);
            \end{tikzpicture}
            \caption{An equivalent system.}
            \label{fig:partRotb}
        \end{subfigure}
        \caption{Diatomic rotation.}
        \label{fig:partRot}
    \end{figure}
    \begin{itemize}
        \item To simplify the problem, replace the two particles rotating about the center of mass with one particle of reduced mass $\mu$ rotating about the center of mass with lever arm $r_0$.
    \end{itemize}
    \item Classically, the kinetic energy of the translational motion is
    \begin{equation*}
        T = \frac{L^2}{2I}
    \end{equation*}
    where $I=\mu r_0^2$ and $L=p\times r_0=pr_0$ (for this kind of rotation; see Figure \ref{fig:partRotb}).
    \item To further talk about this problem, we should introduce \textbf{spherical coordinates}.
    \item \textbf{Spherical coordinates}: The coordinate system $(r,\theta,\phi)$ related to the Cartesian coordinates $(x,y,z)$ by
    \begin{align*}
        x &= r\sin\theta\cos\phi&
        y &= r\sin\theta\sin\phi&
        z &= r\cos\theta
    \end{align*}
    \item Classically, we will have
    \begin{equation*}
        H = \frac{1}{2\mu}(p_x^2+p_y^2+p_z^2)+V(x,y,z)
    \end{equation*}
    in Cartesian coordinates.
    \item In spherical coordinates, this becomes
    \begin{equation*}
        H = \frac{1}{2\mu}\left( p_r^2+\frac{L^2}{r^2} \right)+V(r)
    \end{equation*}
    \item Thus, in quantum mechanics, we get
    \begin{align*}
        \hat{H} &= \frac{1}{2\mu}(\hat{p}_x^2+\hat{p}_y^2+\hat{p}_z^2)+V(x,y,z)\\
        &= \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{\hat{L}^2}{r^2} \right)+V(r)
    \end{align*}
    \item Thus, we have in spherical coordinates that
    \begin{equation*}
        \hat{T}\psi = -\frac{\hbar^2}{2\mu}\left[ \frac{1}{r^2}\pdv{r}\left( r^2\pdv{r}\psi \right)+\frac{1}{r^2\sin\theta}\left( \pdv{\theta}\left( \sin\theta\pdv{\theta}\psi \right) \right)+\frac{1}{r^2\sin^2\theta}\pdv[2]{\phi}\psi \right]
    \end{equation*}
    \item 2D rigid rotor:
    \begin{itemize}
        \item Let the system from Figure \ref{fig:partRotb} be confined to rotating in two dimensions.
        \item This simplifies the problem since both the $\pdv*{r}$ and $\pdv*{\theta}$ terms in the kinetic energy operator disappear (since, respectively, the particle is at a fixed distance from the center of mass and it cannot move out of the 2D plane).
        \item Thus, our Schr\"{o}dinger equation for this system is
        \begin{equation*}
            -\frac{\hbar^2}{2\mu r_0^2}\pdv[2]{\phi}\psi(\phi) = E\psi(\phi)
        \end{equation*}
        \item Solution: Let $\psi(\phi)=\e[im\phi]$; then
        \begin{equation*}
            E_m = \frac{\hbar^2m^2}{2\mu r_0^2}
        \end{equation*}
        \begin{itemize}
            \item $m=0,1,2,\dots$ is a new quantum number.
            \item $m$ doesn't go to infinity because $|m|$ is bounded by $l$ (the total angular momentum).
        \end{itemize}
        \item Remembering our original restriction, we have that this math describes the system from Figure \ref{fig:partRota} but confined to rotate in the $xy$ plane with angular momentum in the $z$ direction.
        \begin{itemize}
            \item Thus, for example, the energies of the system from Figure \ref{fig:partRota} are dependent on $m$ and $I=\mu r_0^2$.
        \end{itemize}
        \item Such a system occurs in physical reality when we put the diatomic in an external field, or attach to it a big functional group.
        \item Zero point energy: $m=0$ does not violate the UR since we still have $\Delta L\Delta\theta\geq\hbar/2$ (as everything is still rotating in the sense that we have equal probability of the particle being everywhere [as opposed to more localized/normal rotation with higher values of $m$]).
    \end{itemize}
    \item 3D rigid rotor:
    \begin{itemize}
        \item Assume that the potential energy is zero on the surface of the sphere (so we basically have a particle on a sphere).
        \item Then
        \begin{equation*}
            \hat{H} = \frac{\hat{L}^2}{2\mu r_0^2} = \frac{\hat{L}^2}{2I}
        \end{equation*}
        \item Solving $\hat{H}\psi=E\psi$ asserts that the eigenfunctions of the Hamiltonian are the spherical harmonics $Y_{l m}(\theta,\phi)$.
        \item Energy:
        \begin{equation*}
            E_l = \frac{\hbar^2}{2I}l(l+1)
        \end{equation*}
        where $l=0,1,2,\dots$.
        \item Recall that $m$ corresponds to the projection of angular momentum onto the $z$-axis, so that
        \begin{equation*}
            m = -l,\dots,+l
        \end{equation*}
    \end{itemize}
\end{itemize}



\section{Hydrogen Atom}
\begin{itemize}
    \item \marginnote{10/20:}Microwaves (for food) excite the rotational motion of water molecules.
    \item Spherical harmonics: The solution of $\psi_{lm}(\theta,\phi)=Y_{lm}(\theta,\phi)$, where $l,m$ are quantum numbers.
    \item $E_l=\hbar^2/2I\cdot l(l+1)$ for $l=0,1,2,\dots$.
    \item Form of the spherical harmonics:
    \begin{equation*}
        Y(\theta,\phi) = P_{lm}(\cos\theta)\e[im\phi]
    \end{equation*}
    where $P_{lm}(cos\theta)$ is a polynomial.
    \item The polynomials $P_{lm}(\cos\theta)$ are the associated Legendre polynomials.
    \item When $m=0$, we have the Legendre polynomials.
    \begin{itemize}
        \item The differential equation describing these is
        \begin{equation*}
            \dv{x}\left( (1-x^2)\dv{x}P_l(x)+l(l+1)P_l(x) \right) = 0
        \end{equation*}
        \item The Legendre polynomials converge very quickly to functions on $[-1,1]$.
        \item People map these polynomials onto other domains, too, to solve a variety of problems.
        \item Legendre polynomials have more of their roots at the boundaries --- since the boundary conditions are the most important part of solving a differential equation, it makes sense that accurate representations would sample near the boundary more.
        \item Examples:
        \begin{align*}
            P_0(x) &= 1&
            P_1(x) &= x\\
            P_2(x) &= \frac{3}{2}x^2-\frac{1}{2}&
            P_3(x) &= \frac{5}{2}x^3-\frac{3}{2}x
        \end{align*}
    \end{itemize}
    \item Consider \ce{HCl}.
    \begin{itemize}
        \item For it,
        \begin{equation*}
            \frac{\hbar^2}{2I} = \SI{1.3e-3}{\electronvolt}
        \end{equation*}
        \item Rotational spectral lines may arise from different values of the quantum number $l$.
        \item The molecule vibrates in harmonic oscillation with spacings $\approx\SI{0.1}{\electronvolt}$ (it's not strictly rigid).
        \item Rovibrational spectra includes both forms of movement.
        \begin{itemize}
            \item Very high precision.
            \item Very big in the 90s.
        \end{itemize}
        \item Electronic spectra: A few electron volts.
    \end{itemize}
    \item The hydrogen atom.
    \begin{itemize}
        \item Two generalizations of the 3D rigid rotor combine to treat the hydrogen atom:
        \begin{enumerate}
            \item An addition of the kinetic energy in the radial direction $\hat{r}$.
            \item An addition of the Coulomb potential.
        \end{enumerate}
        \item Schr\"{o}dinger equation:
        \begin{equation*}
            \hat{H}\psi(r,\theta,\phi) = E\psi(r,\theta,\phi)
        \end{equation*}
        where
        \begin{gather*}
            \hat{H} = \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{\hat{L}^2}{r^2} \right)+V(r)\\
            \hat{p}_r^2 = -\frac{\hbar^2}{r^2}\pdv{r}\left( r^2\pdv{r}\psi(r) \right)\\
            V(r) = -\frac{e(eZ)}{4\pi\epsilon_0r}
        \end{gather*}
        \item Because the particle has spherical symmetry (that it, does not depend on $\theta$ or $\phi$), the wave function is separable, that is, it may be written as a product
        \begin{equation*}
            \psi(r,\theta,\phi) = R_n(r)Y_{lm}(\theta,\phi)
        \end{equation*}
        \item Note that there is no analytic solution to the Schr\"{o}dinger equation in Cartesian coordinates --- we need spherical coordinates to take advantage of the spherical symmetry.
        \item Substitution into the Schr\"{o}dinger equation yields
        \begin{equation*}
            \left( \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{\hat{L}^2}{r^2} \right)+V(r) \right)R(r)Y(\theta,\phi) = ER(r)Y(\theta,\phi)
        \end{equation*}
        where $\mu$ is the reduced mass of the electron and the nucleus (which is approximately the mass of the electron).
        \item But since
        \begin{equation*}
            \frac{1}{2\mu r^2}\hat{L}^2Y_{lm}(\theta,\phi) = \frac{l(l+1)}{2\mu r^2}\hbar^2Y_{lm}(\theta,\phi)
        \end{equation*}
        we have that
        \begin{equation*}
            \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{l(l+1)\hbar^2}{r^2}+V(r) \right)R_n(r) = E_nR_n(r)
        \end{equation*}
        \begin{itemize}
            \item We have reduced the three-dimensional case of the hydrogen atom to a one-dimensional differential equation.
        \end{itemize}
    \end{itemize}
\end{itemize}



\section{Hydrogen Atom (cont.)}
\begin{itemize}
    \item \marginnote{10/22:}We want to solve the SE $\hat{H}\psi_n=E\psi_n$ for the hydrogen atom.
    \item To exploit the spherical symmetry of the hydrogen atom, we use $(r,\theta,\phi)$.
    \item Thus, our Hamiltonian is equal to
    \begin{equation*}
        \hat{H} = \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{\hat{L}^2}{r^2} \right)+V(r)
    \end{equation*}
    \begin{itemize}
        \item Since $V$ is just a function of $r$, the wave function is separable: $\psi(r,\theta,\phi)=R_n(r)Y_{l,m}(\theta,\phi)$.
        \item "Whenever you have a separation of variables additively in the Hamiltonian, you have a separation of variables multiplicatively in the wave function."
    \end{itemize}
    \item Substitute the wave function into the Schr\"{o}dinger equation:
    \begin{align*}
        \left( \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{\hat{L}^2}{r^2} \right)+V(r) \right)R_n(r)Y_{lm}(\theta,\phi) &= ER_n(r)Y_{lm}(\theta,\phi)\\
        \left( \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{l(l+1)\hbar^2}{r^2}\right)+V(r) \right)R_n(r)Y_{lm}(\theta,\phi) &= ER_n(r)Y_{lm}(\theta,\phi)\\
        \left( \frac{1}{2\mu}\left( \hat{p}_r^2+\frac{l(l+1)\hbar^2}{r^2} \right)+V(r) \right)R_n(r) &= E_{nl}R_n(r)
    \end{align*}
    \item Noting that we only care about the behavior of the differential equation on $[0,\infty)$, specifically at really large distances, we perform an asymptotic analysis.
    \begin{equation*}
        \lim_{r\to\infty}\left( \frac{1}{2\mu}\hat{p}_r^2 \right)R(r) = ER(r)
    \end{equation*}
    \item Plugging in the value of the momentum operator, we have that
    \begin{align*}
        -\frac{\hbar^2}{2\mu}\dv[2]{r}R(r) &= ER(r)\\
        R(r) &= \e[-\alpha r]
    \end{align*}
    where $\alpha=i\hbar/\sqrt{2\mu E}$.
    \item But we want to multiply $\e[-\alpha r]$ by some type of polynomial. Namely, \textbf{Laguerre polynomials}.
    \begin{itemize}
        \item These polynomials may be found by expanding a power series.
        \item Let
        \begin{equation*}
            L\left( \frac{\alpha r}{n} \right) = \sum_{j=0}^\infty c_j\left( \frac{\alpha r}{n} \right)^j
        \end{equation*}
        \item A recursion relation may be found.
        \item The quantization of energy in the hydrogen atom again arises from the truncation of the polynomials.
    \end{itemize}
    \item Thus, the general solution of $R_{nl}(r)$ is
    \begin{equation*}
        R_{nl}(r) = \left( \frac{\alpha r}{n} \right)^lL_{n+1,2l+1}\left( \frac{\alpha r}{n} \right)\e[-\alpha r/2n]
    \end{equation*}
    \item Energy levels:
    \begin{equation*}
        E_n = -\frac{\mu}{2\hbar^2}\left( \frac{(Ze)e}{4\pi\epsilon} \right)^2\frac{1}{n^2}
    \end{equation*}
    \begin{itemize}
        \item These are the discrete energies of Bohr!
    \end{itemize}
    \item But now we also have the electron's probability distributions, i.e., the ground state wave function is
    \begin{equation*}
        \psi(r,\theta,\phi) = \frac{1}{\sqrt{\pi}}\left( \frac{Z}{a_0} \right)^{3/2}\e[-Zr/a_0]
    \end{equation*}
    \item Thus, we can now take a look at the probabilities.
    \begin{itemize}
        \item The probability is
        \begin{equation*}
            \text{Pr} = |\psi(r,\theta,\phi)|^2r^2\sin\theta
        \end{equation*}
        \item The radial probability is
        \begin{equation*}
            \text{Pr}(r) = r^2|R_{nl}(r)|^2
        \end{equation*}
        \begin{itemize}
            \item The radial probability peaks at $a_0$, the \textbf{Bohr radius}.
            \item Thus, the Bohr radius (the radius of the circular orbit of Bohr's hydrogen electron) is just the most probable distance from the nucleus!
        \end{itemize}
    \end{itemize}
    \item The average distance from the nucleus (the expectation value of $r$) is
    \begin{align*}
        \prb{r} &= \int_0^\infty\text{Pr}(r)r\dd{r}\\
        &= \int_0^\infty\psi^*(r)r\psi(r)\dd{r}
    \end{align*}
    \begin{itemize}
        \item Note we also find that $\prb{\hat{H}}=E$, so there is no uncertainty in the energy.
    \end{itemize}
    \item Note that in some ways, quantum mechanics is more certain than classical mechanics since, for instance, in quantum we know the energy exactly.
\end{itemize}



\section{Chapter 5: The Harmonic Oscillator and the Rigid Rotator --- Two Spectroscopic Models}
\emph{From \textcite{bib:McQuarrieSimon}.}
\begin{itemize}
    \item \marginnote{10/19:}\textbf{Rigid-rotator model}: Two point masses $m_1$ and $m_2$ at fixed distances $r_1$ and $r_2$ from their center of mass.
    \begin{itemize}
        \item Since the vibrational amplitude of a rotating molecule is small compared to its amplitude, this is a good model.
    \end{itemize}
    \item Kinetic energy of the rigid rotator:
    \begin{align*}
        K &= \frac{1}{2}m_1v_1^2+\frac{1}{2}m_2v_2^2\\
        &= \frac{1}{2}(m_1r_1^2+m_2r_2^2)\omega^2\\
        &= \frac{1}{2}I\omega^2
    \end{align*}
    \begin{itemize}
        \item Note that
        \begin{align*}
            I &= m_1r_1^2+m_2r_2^2\\
            &= \mu r^2
        \end{align*}
        (see Problem \ref{prb:5-29}).
    \end{itemize}
    \item It follows that we the two-body problem of the rigid rotator is equivalent to the one-body problem of a single body of mass $\mu$ rotating at a distance $r$ from a fixed center.
    \item Since there are no external forces on the rigid rotator (we're not applying any electric or magnetic fields), the energy of the molecule is solely kinetic (i.e., there is no potential energy term in the Hamiltonian).
    \begin{itemize}
        \item Thus, for a rigid rotator,
        \begin{equation*}
            \hat{H} = \hat{K} = -\frac{\hbar^2}{2\mu}\nabla^2
        \end{equation*}
    \end{itemize}
    \item Since this particle has a natural center of spherical symmetry, we opt for spherical coordinates. However, this necessitates expressing $\nabla^2$ as the following.
    \begin{equation*}
        \nabla^2 = \frac{1}{r^2}\pdv{r}\left( r^2\pdv{r} \right)_{\theta,\phi}+\frac{1}{r^2\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\theta} \right)_{r,\phi}+\frac{1}{r^2\sin^2\theta}\left( \pdv[2]{\phi} \right)_{r,\theta}
    \end{equation*}
    \begin{itemize}
        \item See Problem \ref{prb:5-32} for a derivation.
    \end{itemize}
    \item With respect to the rigid rotator, $r$ is constant. Thus,
    \begin{gather*}
        \nabla^2 = \frac{1}{r^2}\frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\theta} \right)+\frac{1}{r^2}\frac{1}{\sin^2\theta}\pdv[2]{\phi}\\
        \hat{H} = -\frac{\hbar^2}{2I}\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\theta} \right)+\frac{1}{\sin^2\theta}\left( \pdv[2]{\phi} \right) \right]\\
        \hat{L}^2 = -\hbar^2\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\theta} \right)+\frac{1}{\sin^2\theta}\left( \pdv[2]{\phi} \right) \right]
    \end{gather*}
    \begin{itemize}
        \item Note that since both $\theta$ and $\phi$ are unitless, the units of angular momentum for quantum system are $\hbar$.
    \end{itemize}
    \item Rigid-rotator wave functions are customarily denoted by $Y(\theta,\phi)$.
    \item In solving $\hat{H}Y(\theta,\phi)=EY(\theta,\phi)$, it will be useful to multiply the original Schr\"{o}dinger equation by $\sin^2\theta$ and let $\beta=2IE/\hbar^2$ to obtain the partial differential equation
    \begin{equation*}
        \sin\theta\pdv{\theta}\left( \sin\theta\pdv{Y}{\theta} \right)+\pdv[2]{Y}{\phi}+(\beta\sin^2\theta)Y = 0
    \end{equation*}
    \begin{itemize}
        \item The solutions to the above equation are intimately linked to those for the hydrogen atom.
        \item Solving the above equation yields the condition that $\beta=J(J+1)$ for $J=0,1,2,\dots$. Therefore,
        \begin{equation*}
            E_J = \frac{\hbar^2}{2I}J(J+1)
        \end{equation*}
        for $J=0,1,2,\dots$.
        \begin{itemize}
            \item Each energy level has a degeneracy $g_J=2J+1$ as well.
        \end{itemize}
    \end{itemize}
    \item Once again, electromagnetic radiation can cause a rigid rotator to undergo transitions from one state to another subject to the selection rules that only transitions between adjacent states are allowed and the molecule must possess a permanent dipole moment.
    \item As before, we can calculate
    \begin{equation*}
        \Delta E = \frac{h^2}{4\pi^2I}(J+1)
    \end{equation*}
    and the frequencies at which absorption transitions occur are
    \begin{equation*}
        \nu = \frac{h}{4\pi^2I}(J+1)
    \end{equation*}
    for $J=0,1,2,\dots$.
    \item It follows from reduced mass, bond length, and moment of inertia data that the frequencies typically lie in the microwave region.
    \item \textbf{Microwave spectroscopy}: The direct study of rotational transitions.
    \item \textbf{Rotational constant} (of a molecule): The following quantity. \emph{Given by}
    \begin{equation*}
        B = \frac{h}{8\pi^2I}
    \end{equation*}
    \begin{itemize}
        \item We often write the absorption frequencies as $\nu=2B(J+1)$.
    \end{itemize}
    \item The spacing of lines in a microwave spectrum is $2B$.
    \item Like IR spectroscopy can be used to determine the force constants of molecular attractions in diatomics, microwave spectroscopy can be used to determine the bond lengths of diatomics.
\end{itemize}


\subsection*{Problems}
\begin{enumerate}[label={\textbf{5-\arabic*.}},ref={5-\arabic*}]
    \setcounter{enumi}{23}
    \item \label{prb:5-24}\marginnote{12/1:}There are a number of general relations between the Hermite polynomials and their derivatives (which we will not derive). Some of these are
    \begin{gather*}
        \dv{H_v(\xi)}{\xi} = 2\xi H_v(\xi)-H_{v+1}(\xi)\\
        H_{v+1}(\xi)-2\xi H_v(\xi)+2vH_{v-1}(\xi) = 0\\
        \dv{H_v(\xi)}{\xi} = 2vH_{v-1}(\xi)
    \end{gather*}
    Such connecting relations are called \textbf{recursion formulas}. Verify these formulas explicitly using the first few Hermite polynomials given in Table \ref{tab:hermitePolynomials}.
    \setcounter{enumi}{28}
    \item \label{prb:5-29}\marginnote{10/19:}Show that the moment of inertia for a rigid rotator can be written as $I=\mu r^2$ where $r=r_1+r_2$ (the fixed separation of the two masses) and $\mu$ is the reduced mass.
    \begin{proof}[Answer]
        First, note that $m_1r_1=m_2r_2$ for such a rotation about the center of mass. Then
        \begingroup
        \allowdisplaybreaks
        \begin{align*}
            I &= \mu r^2\\
            &= \frac{m_1m_2}{m_1+m_2}(r_1+r_2)^2\\
            &= \frac{m_1m_2r_1}{m_1r_1+m_2r_1}(r_1+r_2)^2\\
            &= \frac{m_1m_2r_1}{m_2r_2+m_2r_1}(r_1+r_2)^2\\
            &= \frac{m_1r_1}{r_2+r_1}(r_1+r_2)^2\\
            &= m_1r_1(r_1+r_2)\\
            &= m_1r_1^2+m_1r_1r_2\\
            &= m_1r_1^2+m_2r_2^2
        \end{align*}
        \endgroup
        as desired.
    \end{proof}
    \item \label{prb:5-30}Consider the transformation from Cartesian coordinates to plane polar coordinates where
    \begin{align*}
        x &= r\cos\theta&
            r &= \sqrt{x^2+y^2}\\
        y &= r\sin\theta&
            \theta &= \tan^{-1}\left( \frac{y}{x} \right)
    \end{align*}
    If a function $f(r,\theta)$ depends upon the polar coordinates $r$ and $\theta$, then the chain rule of partial differentiation says that\footnote{Note that the subscript means that the subscripted variable is held constant.}
    \begin{equation*}
        \left( \pdv{f}{x} \right)_y = \left( \pdv{f}{r} \right)_\theta\left( \pdv{r}{x} \right)_y+\left( \pdv{f}{\theta} \right)_r\left( \pdv{\theta}{x} \right)_y
    \end{equation*}
    and that
    \begin{equation*}
        \left( \pdv{f}{y} \right)_x = \left( \pdv{f}{r} \right)_\theta\left( \pdv{r}{y} \right)_x+\left( \pdv{f}{\theta} \right)_r\left( \pdv{\theta}{y} \right)_x
    \end{equation*}
    For simplicity, we will assume that $r$ is constant so that we can ignore terms involving derivatives with respect to $r$. In other words, we will consider a particle that is constrained to move on the circumference of a circle. This system is sometimes called a \textbf{particle on a ring}. Using the above equations, show that
    \begin{align*}
        \left( \pdv{f}{x} \right)_y &= -\frac{\sin\theta}{r}\left( \pdv{f}{\theta} \right)_r&
        \left( \pdv{f}{y} \right)_x &= \frac{\cos\theta}{r}\left( \pdv{f}{\theta} \right)_r
    \end{align*}
    for $r$ fixed. Now apply the above equations again to show that
    \begin{align*}
        \left( \pdv[2]{f}{x} \right)_y &= \left[ \pdv{x}\left( \pdv{f}{x} \right)_y \right]\\
        &= \left[ \pdv{\theta}\left( \pdv{f}{x} \right)_y \right]_r\left( \pdv{\theta}{x} \right)_y\\
        &= \left\{ \pdv{\theta}\left[ -\frac{\sin\theta}{r}\left( \pdv{f}{\theta} \right)_r \right] \right\}_r\left( -\frac{\sin\theta}{r} \right)\\
        &= \frac{\sin\theta\cos\theta}{r^2}\left( \pdv{f}{\theta} \right)_r+\frac{\sin^2\theta}{r^2}\left( \pdv[2]{f}{\theta} \right)_r
    \end{align*}
    for $r$ fixed. Similarly, show that
    \begin{equation*}
        \left( \pdv[2]{f}{y} \right)_x = -\frac{\sin\theta\cos\theta}{r^2}\left( \pdv{f}{\theta} \right)_r+\frac{\cos^2\theta}{r^2}\left( \pdv[2]{f}{\theta} \right)_r
    \end{equation*}
    and that
    \begin{equation*}
        \nabla^2f = \pdv[2]{f}{x}+\pdv[2]{f}{y} \longrightarrow \frac{1}{r^2}\left( \pdv[2]{f}{\theta} \right)_r
    \end{equation*}
    both for $r$ fixed. Now show that the Schr\"{o}dinger equation for a particle of mass $m$ constrained to move on a circle of radius $r$ (see Problem \ref{prb:3-28}) is
    \begin{equation*}
        -\frac{\hbar^2}{2I}\pdv[2]{\psi(\theta)}{\theta} = E\psi(\theta)
    \end{equation*}
    where $I=mr^2$ is the moment of inertia and $0\leq\theta\leq 2\pi$.
    \begin{proof}[Answer]
        We have that
        \begin{align*}
            \left( \pdv{\theta}{x} \right)_y &= \pdv{x}\left( \tan^{-1}\left( \frac{y}{x} \right) \right)&
                \left( \pdv{\theta}{y} \right)_y &= \pdv{y}\left( \tan^{-1}\left( \frac{y}{x} \right) \right)\\
            &= \frac{1}{\left( \frac{y}{x} \right)^2+1}\cdot y\cdot -\frac{1}{x^2}&
                &= \frac{1}{\left( \frac{y}{x} \right)^2+1}\cdot\frac{1}{x}\\
            &= -\frac{y}{x^2+y^2}&
                &= \frac{x}{x^2+y^2}\\
            &= -\frac{r\sin\theta}{r^2}&
                &= \frac{r\cos\theta}{r^2}\\
            &= -\frac{\sin\theta}{r}&
                &= \frac{\cos\theta}{r}
        \end{align*}
        This combined with the fact that $r$ is constant yields
        \begin{align*}
            \left( \pdv{f}{x} \right)_y &= \left( \pdv{f}{r} \right)_\theta\left( \pdv{r}{x} \right)_y+\left( \pdv{f}{\theta} \right)_r\left( \pdv{\theta}{x} \right)_y&
                \left( \pdv{f}{y} \right)_x &= \left( \pdv{f}{r} \right)_\theta\left( \pdv{r}{y} \right)_x+\left( \pdv{f}{\theta} \right)_r\left( \pdv{\theta}{y} \right)_x\\
            &= 0\cdot\left( \pdv{r}{x} \right)_y+\left( \pdv{f}{\theta} \right)_r\cdot -\frac{\sin\theta}{r}&
                &= 0\cdot\left( \pdv{r}{y} \right)_x+\left( \pdv{f}{\theta} \right)_r\frac{\cos\theta}{r}\\
            &= -\frac{\sin\theta}{r}\left( \pdv{f}{\theta} \right)_r&
                &= \frac{\cos\theta}{r}\left( \pdv{f}{\theta} \right)_r
        \end{align*}
        as desired.\par
        The rest of the differential equation derivations follow fairly easy with simple calculus rules.\par
        As for the Schr\"{o}dinger equation, we know that there is no potential field for this particle. Additionally, although we could choose to express $\psi$ as a function of $x,y$, owing to the radial symmetry, we choose to express it as a function of $r,\theta$. In fact, we need only express $\psi$ as a function of $\theta$ since $r$ is invariant throughout the chosen free space. Thus, expanding from the general form, we get
        \begin{align*}
            \hat{H}\psi(\theta) &= E\psi(\theta)\\
            \left( -\frac{\hbar^2}{2m}\nabla^2 \right)\psi(\theta) &= E\psi(\theta)\\
            -\frac{\hbar^2}{2m}\cdot\frac{1}{r^2}\left( \pdv[2]{\psi(\theta)}{\theta} \right)_r &= E\psi(\theta)\\
            -\frac{\hbar^2}{2I}\pdv[2]{\psi(\theta)}{\theta} &= E\psi(\theta)
        \end{align*}
        as desired.
    \end{proof}
    \item \label{prb:5-31}Generalize Problem \ref{prb:5-30} to the case of a particle moving in a plane under the influence of a central force; in other words, convert
    \begin{equation*}
        \nabla^2 = \pdv[2]{x}+\pdv[2]{y}
    \end{equation*}
    to plane polar coordinates, this time without assuming that $r$ is constant. Use the method of separation of variables to separate the equation for this problem. Solve the angular equation.
    \begin{proof}[Answer]
        We have
        \begin{align*}
            \pdv{r}{x} &= \frac{1}{2\sqrt{x^2+y^2}}\cdot 2x&
                \pdv{r}{y} &= \frac{1}{2\sqrt{x^2+y^2}}\cdot 2y\\
            &= \frac{x}{r}&
                &= \frac{y}{r}\\
            &= \cos\theta&
                &= \sin\theta
        \end{align*}
        Thus
        \begin{align*}
            \pdv{f}{x} &= \cos\theta\pdv{f}{r}-\frac{\sin\theta}{r}\pdv{f}{\theta}&
            \pdv{f}{y} &= \sin\theta\pdv{f}{r}+\frac{\cos\theta}{r}\pdv{f}{\theta}
        \end{align*}
        Also note that
        \begin{align*}
            \pdv[2]{f}{x} &= \pdv{x}\left( \pdv{f}{x} \right)\\
            &= \pdv{\theta}\left( \pdv{f}{x} \right)\pdv{\theta}{x}+\pdv{r}\left( \pdv{f}{x} \right)\pdv{r}{x}
        \end{align*}
        The rest of the expansions are routine, leading to
        \begin{equation*}
            \nabla^2f = \pdv[2]{u}{r}+\frac{1}{r}\pdv{u}{r}+\frac{1}{r^2}\pdv[2]{u}{\theta}
        \end{equation*}
    \end{proof}
    \item \label{prb:5-32}Using Problems \ref{prb:5-30} and \ref{prb:5-31} as a guide, convert $\nabla^2$ from three-dimensional Cartesian coordinates to spherical coordinates.
\end{enumerate}



\section{Chapter 6: The Hydrogen Atom}
\emph{From \textcite{bib:McQuarrieSimon}.}
\begin{itemize}
    \item \marginnote{10/29:}Idealize the hydrogen atom to a proton fixed at the origin and an electron of mass $m_e$ interacting with the proton through a Coulombic potential
    \begin{equation*}
        V(r) = -\frac{e^2}{4\pi\epsilon_0r}
    \end{equation*}
    \item Because of the spherical geometry of the model, we opt for spherical coordinates.
    \item The appropriate Hamiltonian is thus
    \begin{equation*}
        \hat{H} = -\frac{\hbar}{2m_e}\nabla^2-\frac{e^2}{4\pi\epsilon_0r}
    \end{equation*}
    \item Expressing the Laplacian in spherical coordinates gives us
    \begin{equation*}
        \left\{ -\frac{\hbar^2}{2m_e}\left[ \frac{1}{r^2}\pdv{r}\left( r^2\pdv{\psi}{r} \right)+\frac{1}{r^2\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\psi}{\theta} \right)+\frac{1}{r^2\sin^2\theta}\pdv[2]{\psi}{\phi} \right]-\frac{e^2}{4\pi\epsilon_0r} \right\}\psi(r,\theta,\phi) = E\psi(r,\theta,\phi)
    \end{equation*}
    as the Schr\"{o}dinger equation for the hydrogen atom.
    \item Multiply through by $2m_er^2$ and move all terms over to the lefthand side of the equality to get the above equation into the following separable form.
    \begin{equation*}
        \left\{ -\hbar^2\left( \pdv{r}r^2\pdv{\psi}{r} \right)-\hbar^2\left[ \frac{1}{\sin\theta}\left( \pdv{\theta}\sin\theta\pdv{\psi}{\theta} \right)+\frac{1}{\sin^2\theta}\pdv[2]{\psi}{\phi} \right]-2m_er^2\left[ \frac{e^2}{4\pi\epsilon_0r}+E \right] \right\}\psi(r,\theta,\phi) = 0
    \end{equation*}
    \item Since all the $\theta,\phi$ dependence occurs within the large square brackets, assume that
    \begin{equation*}
        \psi(r,\theta,\phi) = R(r)Y(\theta,\phi)
    \end{equation*}
    \item Substituting into the differential equation and separating terms (dividing by $R(r)Y(\theta,\phi)$) yields
    \begin{equation*}
        -\frac{\hbar^2}{R(r)}\left[ \dv{r}\left( r^2\dv{R}{r} \right)+\frac{2m_er^2}{\hbar^2}\left( \frac{e^2}{4\pi\epsilon_0r}+E \right)R(r) \right]-\frac{\hbar^2}{Y(\theta,\phi)}\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{Y}{\theta} \right)+\frac{1}{\sin^2\theta}\pdv[2]{Y}{\phi} \right] = 0
    \end{equation*}
    \item All terms containing $r$ are in the first set of square brackets, and all terms containing $\theta,\phi$ are in the second set of square brackets. Since the above equation must hold for all $r,\theta,\phi$, and $r,\theta,\phi$ are independent variables, we must have that varying $r$ while holding $\theta,\phi$ constant does not change the $r$ term of the above equation (as this would violate the equality). Similarly, changing $\theta,\phi$ must not change the $\theta,\phi$ term. Therefore, we may let
    \begin{gather*}
        -\beta = -\frac{1}{R(r)}\left[ \dv{r}\left( r^2\dv{R}{r} \right)+\frac{2m_er^2}{\hbar^2}\left( \frac{e^2}{4\pi\epsilon_0r}+E \right)R(r) \right]\\
        \beta = -\frac{1}{Y(\theta,\phi)}\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{Y}{\theta} \right)+\frac{1}{\sin^2\theta}\pdv[2]{Y}{\phi} \right]
    \end{gather*}
    where $\beta$ is a \textbf{separation constant} into which we have incorporated $\hbar^2$.
    \item \textbf{Radial equation}: The first equation above.
    \item \textbf{Angular equation}: The second equation above after having been multiplied by $\sin^2\theta Y(\theta,\phi)$. \emph{Given by}
    \begin{equation*}
        \sin\theta\pdv{\theta}\left( \sin\theta\pdv{Y}{\theta}+\pdv[2]{Y}{\phi}+\beta\sin^2\theta Y \right) = 0
    \end{equation*}
    \begin{itemize}
        \item Note that this is identical to the partial differential equation derived in Chapter 5 for the rigid rotator.
        \item Thus, the angular parts of hydrogen atomic orbitals are also rigid-rotator wave functions.
    \end{itemize}
    \item \textbf{Spherical harmonic}: A wave function of the rigid rotator.
    \item We now solve the angular equation.
    \item Invoking separation of variables again with $Y(\theta,\phi)=\Theta(\theta)\Phi(\phi)$ yields
    \begin{equation*}
        \frac{\sin\theta}{\Theta(\theta)}\dv{\theta}\left( \sin\theta\dv{\Theta}{\theta} \right)+\beta\sin\theta+\frac{1}{\Phi(\phi)}\dv[2]{\Phi}{\phi} = 0
    \end{equation*}
    \item We now introduce a new separation constant (written in a form that predicts future algebraic manipulations) to get
    \begin{gather*}
        \frac{\sin\theta}{\Theta(\theta)}\dv{\theta}\left( \sin\theta\dv{\Theta}{\theta} \right)+\beta\sin^2\theta = m^2\\
        \frac{1}{\Phi(\phi)}\dv[2]{\Phi}{\phi} = -m^2
    \end{gather*}
    \item Solving the bottom equation above gives solutions
    \begin{align*}
        \Phi(\phi) &= A_m\e[im\phi]&
        \Phi(\phi) &= A_{-m}\e[-im\phi]
    \end{align*}
    where the constants $A_m,A_{-m}$ may depend on the value of $m$.
    \item From the domain of $\phi$, we have the boundary condition $\Phi(\phi+2\pi)=\Phi(\phi)$.
    \item This yields
    \begin{align*}
        A_m\e[im(\phi+2\pi)] &= A_m\e[im\phi]&
            A_{-m}\e[-im(\phi+2\pi)] &= A_{-m}\e[-im\phi]\\
        \e[i2\pi m] &= 1&
            \e[-i2\pi m] &= 1
    \end{align*}
    from which it follows that we must have
    \begin{align*}
        1+0i &= 1\\
        &= \e[\pm i2\pi m]\\
        &= \cos(2\pi m)\pm i\sin(2\pi m)
    \end{align*}
    \item Therefore, $m=0,\pm 1,\pm 2,\dots$.
    \item It follows that we can express the solutions to the bottom angular equation in the more compact form
    \begin{equation*}
        \Phi_m(\phi) = A_m\e[im\phi]
    \end{equation*}
    where $m=0,\pm 1,\pm 2,\dots$.
    \item Normalizing gives us the coefficient $A_m$:
    \begin{align*}
        1 &= \int_0^{2\pi}\Phi_m^*(\phi)\Phi_m(\phi)\dd{\phi}\\
        &= \int_0^{2\pi}(A_m\e[-im\phi])(A_m\e[im\phi])\dd{\phi}\\
        &= A_m^2\int_0^{2\pi}1\dd{\phi}\\
        A_m &= \frac{1}{\sqrt{2\pi}}
    \end{align*}
    \item Thus, the normalized wave functions of $\Phi(\phi)$ are
    \begin{equation*}
        \Phi(\phi) = \frac{1}{\sqrt{2\pi}}\e[im\phi]
    \end{equation*}
    for $m=0,\pm 1,\pm 2,\dots$.
    \item We now direct our attention to solving the differential equation containing $\Theta(\theta)$.
    \item Let $x=\cos\theta$ and $P(x)=\Theta(\theta)$. Substituting, we have (see Problem \ref{prb:6-2})
    \begin{equation*}
        (1-x^2)\dv[2]{P}{x}-2x\dv{P}{x}+\left[ \beta-\frac{m^2}{1-x^2} \right]P(x) = 0
    \end{equation*}
    where $m=0,\pm 1,\pm 2,\dots$.
    \begin{itemize}
        \item Note that since $x=\cos\theta$, $0\leq\theta\leq\pi$, the range of $x$ is $-1\leq x\leq 1$.
    \end{itemize}
    \item Solving the above equation gives \textbf{Legendre's equation}.
    \item \textbf{Legendre's equation}: The following equation. \emph{Given by}
    \begin{equation*}
        (1-x^2)\dv[2]{P}{x}-2x\dv{P}{x}+\left[ l(l+1)-\frac{m^2}{1-x^2} \right]P(x) = 0
    \end{equation*}
    \begin{itemize}
        \item The key difference between Legendre's equation and the previous one is that solving the previous one finds $\beta$ equal to $l(l+1)$ with $l=0,1,2,\dots$ and $|m|\leq l$ if the solutions are to remain finite.
    \end{itemize}
    \item \textbf{Legendre polynomials}: The solutions to Legendre's equation when $m=0$. \emph{Denoted by} $\bm{P_l(x)}$.
    \begin{table}[h!]
        \centering
        \small
        \renewcommand{\arraystretch}{1.4}
        \begin{tabular}{l}
            \toprule
            $P_0(x)=1$\\
            $P_1(x)=x$\\
            $P_2(x)=\frac{1}{2}(3x^2-1)$\\
            $P_3(x)=\frac{1}{2}(5x^3-3x)$\\
            $P_4(x)=\frac{1}{8}(35x^4-30x^2+3)$\\
            \bottomrule
        \end{tabular}
        \caption{The first few Legendre polynomials.}
        \label{tab:legendrePolynomials}
    \end{table}
    \begin{itemize}
        \item Notice that $P_l(x)$ is an even function if $l$ is even and an odd function if $l$ is odd.
        \item The factors in front of $P_l(x)$ are chosen such that $P_l(1)=1$.
        \item The Legendre polynomials are orthogonal: If $l\neq n$, then
        \begin{equation*}
            \int_{-1}^1P_l(x)P_n(x)\dd{x} = 0
        \end{equation*}
        \item We have that
        \begin{equation*}
            \int_{-1}^1[P_l(x)]^2\dd{x} = \frac{2}{2l+1}
        \end{equation*}
        for each $l$, so the normalization constant of $P_l(x)$ is
        \begin{equation*}
            \sqrt{\frac{2l+l}{2}}
        \end{equation*}
    \end{itemize}
    \item \textbf{Associated Legendre functions}: The solutions to Legendre's equation when $m\neq 0$. \emph{Denoted by} $\bm{P_l^{|m|}(x)}$. \emph{Given by}
    \begin{equation*}
        P_l^{|m|}(x) = (1-x^2)^{|m|/2}\dv[|m|]{x}P_l(x)
    \end{equation*}
    \begin{itemize}
        \item Note that only the magnitude of $m$ is present because $m^2$ is the only form of $m$ present in Legendre's equation.
    \end{itemize}
    \begin{table}[h!]\marginnote{10/30:}
        \centering
        \small
        \renewcommand{\arraystretch}{1.4}
        \begin{tabular}{lll}
            \toprule
            \textbf{Function} & \textbf{$\bm{x}$-coordinates} & \textbf{$\bm{\theta}$-coordinates}\\
            \midrule
            $P_0^0(x)$ & $1$ & $1$\\
            $P_1^0(x)$ & $x$ & $\cos\theta$\\
            $P_1^1(x)$ & $\sqrt{1-x^2}$ & $\sin\theta$\\
            $P_2^0(x)$ & $\frac{1}{2}(3x^2-1)$ & $\frac{1}{2}(3\cos^2\theta-1)$\\
            $P_2^1(x)$ & $3x\sqrt{1-x^2}$ & $3\cos\theta\sin\theta$\\
            $P_2^2(x)$ & $3(1-x^2)$ & $3\sin^2\theta$\\
            $P_3^0(x)$ & $\frac{1}{2}(5x^2-3x)$ & $\frac{1}{2}(5\cos^2\theta-3\cos\theta)$\\
            $P_3^1(x)$ & $\frac{3}{2}(5x^2-1)\sqrt{1-x^2}$ & $\frac{3}{2}(5\cos^2\theta-1)\sin\theta$\\
            $P_3^2(x)$ & $15x(1-x^2)$ & $15\cos\theta\sin^2\theta$\\
            $P_3^3(x)$ & $15\sqrt{1-x^2}$ & $15\sin^2\theta$\\
            \bottomrule
        \end{tabular}
        \caption{The first few associated Legendre functions.}
        \label{tab:associatedLegendreFunctions}
    \end{table}
    \item Relating the normalization and orthogonality conditions back to the variable of interest ($\theta$), we have (for the Legendre polynomials) that
    \begin{equation*}
        \int_0^\pi P_l(\cos\theta)P_n(\cos\theta)\sin\theta\dd{\theta} = \frac{2\delta_{ln}}{2l+1}
    \end{equation*}
    \item Similarly, for the associated Legendre functions, we have that
    \begin{equation*}
        \int_0^\pi P_l^{|m|}(\cos\theta)P_n^{|m|}(\cos\theta)\sin\theta\dd{\theta} = \frac{2}{2l+1}\frac{(l+|m|)!}{(l-|m|)!}\delta_{ln}
    \end{equation*}
    \begin{itemize}
        \item Thus, the normalization constant for the associated Legendre functions is
        \begin{equation*}
            N_{lm} = \sqrt{\frac{2l+1}{2}\frac{(l-|m|)!}{(l+|m|)!}}
        \end{equation*}
    \end{itemize}
    \item Thus, the normalized wave functions of $\Theta(\theta)$ are
    \begin{equation*}
        \Theta(\theta) = N_{lm}P_l^{|m|}(\cos\theta)
    \end{equation*}
    for $l=0,1,2,\dots$ and $m=0,\pm 1,\dots,\pm l$.
    \item \textbf{Spherical harmonics}: The orthonormal set of angular wave functions of the hydrogen atom. \emph{Given by}
    \begin{equation*}
        Y_l^m(\theta,\phi) = \Theta(\theta)\Phi(\phi)
        = \sqrt{\frac{2l+1}{4\pi}\frac{(l-|m|)!}{(l+|m|)!}}P_l^{|m|}(\cos\theta)\e[im\phi]
    \end{equation*}
    \begin{itemize}
        \item Note that $l=0,1,2,\dots$ and $m=0,\pm 1,\dots,\pm l$.
        \item The orthonormality condition:
        \begin{equation*}
            \int_0^\pi\dd{\theta}\sin\theta\int_0^{2\pi}\dd{\phi}Y_l^m(\theta,\phi)^*Y_n^k(\theta,\phi) = \delta_{ln}\delta_{mk}
        \end{equation*}
    \end{itemize}
    \begin{table}[h!]
        \centering
        \small
        \renewcommand{\arraystretch}{2.5}
        \begin{tabular}{ll}
            \toprule
            $Y_0^0=\dfrac{1}{\sqrt{4\pi}}$ & $Y_2^1=\sqrt{\dfrac{15}{8\pi}}\sin\theta\cos\theta\e[i\phi]$\\
            $Y_1^0=\sqrt{\dfrac{3}{4\pi}}\cos\theta$ & $Y_2^{-1}=\sqrt{\dfrac{15}{8\pi}}\sin\theta\cos\theta\e[-i\phi]$\\
            $Y_1^1=\sqrt{\dfrac{3}{8\pi}}\sin\theta\e[i\phi]$ & $Y_2^2=\sqrt{\dfrac{15}{32\pi}}\sin^2\theta\e[2i\phi]$\\
            $Y_1^{-1}=\sqrt{\dfrac{3}{8\pi}}\sin\theta\e[-i\phi]$ & $Y_2^{-2}=\sqrt{\dfrac{15}{32\pi}}\sin^2\theta\e[-2i\phi]$\\
            $Y_2^0=\sqrt{\dfrac{5}{16\pi}}(3\cos^2\theta-1)$ & \\
            \bottomrule
        \end{tabular}
        \caption{The first few spherical harmonics.}
        \label{tab:sphericalHarmonics}
    \end{table}
    \item \marginnote{10/31:}Recalling the definition of $\hat{L}^2$ from Chapter 5, we have that
    \begin{align*}
        \beta &= -\frac{1}{Y_l^m(\theta,\phi)}\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{Y_l^m(\theta,\phi)}{\theta} \right)+\frac{1}{\sin^2\theta}\pdv[2]{Y_l^m(\theta,\phi)}{\phi} \right]\\
        \hbar^2l(l+1)Y_l^m(\theta,\phi) &= -\hbar^2\left[ \frac{1}{\sin\theta}\pdv{\theta}\left( \sin\theta\pdv{\theta} \right)+\frac{1}{\sin^2\theta}\pdv[2]{\phi} \right]Y_l^m(\theta,\phi)\\
        \hat{L}^2Y_l^m(\theta,\phi) &= \hbar^2l(l+1)Y_l^m(\theta,\phi)
    \end{align*}
    \begin{itemize}
        \item Implication: The spherical harmonics are eigenfunctions of $\hat{L}^2$.
        \item Additionally, the square of the angular momentum $L^2$ can only have the values given by $\hbar^2l(l+1)$ for $l=0,1,2,\dots$.
        \item We can also adapt the above to reflect energy with $\hat{H}=\hat{L}^2/2I$.
    \end{itemize}
    \item \marginnote{11/1:}We can apply partial differentiation to convert the expressions for $\hat{L}_x,\hat{L}_y,\hat{L}_z$ in Table \ref{tab:observablesOperators} to spherical coordinates.
    \begin{itemize}
        \item Doing so yields complex expressions for $\hat{L}_x$ and $\hat{L}_y$, but a simple expression for $\hat{L}_z$:
        \begin{equation*}
            \hat{L}_z = -i\hbar\dv{\phi}
        \end{equation*}
    \end{itemize}
    \item The spherical harmonics are eigenfunctions of $\hat{L}_z$:
    \begin{align*}
        \hat{L}_zY_l^m(\theta,\phi) &= -i\hbar\dv{\phi}(N_{lm}P_l^{|m|}(\cos\theta)\e[im\phi])\\
        &= -i\hbar N_{lm}P_l^{|m|}(\cos\theta)\dv{\phi}(\e[im\phi])\\
        &= m\hbar N_{lm}P_l^{|m|}(\cos\theta)\e[im\phi]\\
        &= m\hbar Y_l^m(\theta,\phi)
    \end{align*}
    \begin{itemize}
        \item It follows that $L_z$ is quantized by $m\hbar$.
        \item This implies that $\hbar$ is a fundamental measure of the angular momentum of a quantum mechanical system.
        \item Note that the spherical harmonics are not eigenfunctions of $\hat{L}_x$ or $\hat{L}_y$.
    \end{itemize}
    \item The last two results imply that we can determine precise values of $L^2$ and $L_z$ simultaneously, and that $\hat{L}^2,\hat{L}_z$ commute.
    \item Proving that $|m|<l$.
    \begin{itemize}
        \item We have that
        \begin{align*}
            \hat{L}_z^2Y_l^m(\theta,\phi) &= m^2\hbar^2Y_l^m(\theta,\phi)&
            \hat{L}^2Y_l^m(\theta,\phi) &= l(l+1)\hbar^2Y_l^m(\theta,\phi)
        \end{align*}
        \item Subtracting and substituting $\hat{L}^2=\hat{L}_x^2+\hat{L}_y^2+\hat{L}_z^2$ yields
        \begin{align*}
            (\hat{L}^2-\hat{L}_z^2)Y_l^m(\theta,\phi) &= [l(l+1)-m^2]\hbar^2Y_l^m(\theta,\phi)\\
            (\hat{L}_x^2+\hat{L}_y^2)Y_l^m(\theta,\phi) &= [l(l+1)-m^2]\hbar^2Y_l^m(\theta,\phi)
        \end{align*}
        \item It follows since $m,l\in\Z$ that
        \begin{align*}
            0 &\leq L_x^2+L_y^2 = [l(l+1)-m^2]\hbar^2\\
            0 &\leq l(l+1)-m^2\\
            m^2 &\leq l(l+1)\\
            |m| &\leq l
        \end{align*}
        as desired.
    \end{itemize}
    \item Consider the case $l=1$.
    \begin{itemize}
        \item In this case, $L^2=2\hbar^2$ and $L_z=-\hbar,0,\hbar$.
        \item It follows from the first condition that $|L|=\sqrt{L^2}=\sqrt{2}\hbar$, i.e., the magnitude of the angular momentum vector is $\sqrt{2}\hbar$.
        \item Since $L$ and $L_z$ have differing magnitudes, their vectors cannot point in the same direction.
        \item Since $\hat{L}^2$ commutes with each of $\hat{L}_x,\hat{L}_y,\hat{L}_z$ individually but the three directional components don't commute among themselves, it is possible to measure the values of $L^2$ and any directional component simultaneously to any precision, but this leads to complete uncertainty in the other two directional components.
        \item This combined with the fact that $\prb{L_x}=\prb{L_y}=0$ leads to this nice interpretation: Imagine $\mathbf{L}$ processing around a cone whose axis is the $z$-axis. We know that the height of $\mathbf{L}$ (its $z$-coordinate) never changes, but the $x,y$-coordinates could be anywhere.
    \end{itemize}
    \item Solving the radial equation.
    \begin{itemize}
        \item Energy is quantized by
        \begin{equation*}
            E_n = -\frac{m_ee^4}{8\epsilon_0^2h^2n^2}
            = -\frac{m_ee^4}{32\pi^2\epsilon_0^2\hbar^2n^2}
            = -\frac{e^2}{8\pi\epsilon_0a_0n^2}
        \end{equation*}
        for $n=1,2,\dots$ where $a_0$ is the Bohr radius.
        \begin{itemize}
            \item Note that these energies are identical to those predicted by the Bohr model, even though the electron here is not restricted to sharply defined circular orbits but rather is described by $\psi(r,\theta,\phi)$.
        \end{itemize}
        \item We also have the condition
        \begin{equation*}
            0 \leq l \leq n-1
        \end{equation*}
        on the quantum number $n$.
        \item Thus, the radial wave functions depend on the quantum numbers $n$ and $l$ as follows.
        \begin{equation*}
            R_{nl}(r) = -\sqrt{\frac{(n-l-1)!}{2n[(n+l)!]^3}}\left( \frac{2}{na_0} \right)^{l+3/2}r^l\e[-r/na_0]L_{n+l}^{2l+1}\left( \frac{2r}{na_0} \right)
        \end{equation*}
        \begin{itemize}
            \item The $L_{n+l}^{2l+1}(2r/na_0)$ are the \textbf{associated Laguerre polynomials}.
            \begin{table}[h!]\marginnote{11/2:}
                \centering
                \small
                \renewcommand{\arraystretch}{1.4}
                \begin{tabular}{lll}
                    \toprule
                    $n=1$ & $l=0$ & $L_1^1(x)=-1$\\
                    \hline
                    $n=2$ & $l=0$ & $L_2^1(x)=-2!(2-x)$\\
                          & $l=1$ & $L_3^3(x)=-3!$\\
                    \hline
                    $n=3$ & $l=0$ & $L_3^1(x)=-3!(3-3x+\frac{1}{2}x^2)$\\
                          & $l=1$ & $L_4^3(x)=-4!(4-x)$\\
                          & $l=2$ & $L_5^5(x)=-5!$\\
                    \hline
                    $n=4$ & $l=0$ & $L_4^1(x)=-4!(4-6x+2x^2-\frac{1}{6}x^3)$\\
                          & $l=1$ & $L_5^3(x)=-5!(10-5x+\frac{1}{2}x^2)$\\
                          & $l=2$ & $L_6^5(x)=-6!(6-x)$\\
                          & $l=3$ & $L_7^7(x)=-7!$\\
                    \bottomrule
                \end{tabular}
                \caption{The first few associated Laguerre polynomials.}
                \label{tab:associatedLaguerrePolynomials}
            \end{table}
            \item The combinatorial factors in Table \ref{tab:associatedLaguerrePolynomials} just guarantee that the $R_{nl}(r)$ are normalized.
        \end{itemize}
    \end{itemize}
    \item We can now construct the complete hydrogen atomic wave functions as
    \begin{equation*}
        \psi_{nlm}(r,\theta,\phi) = R_{nl}(r)Y_l^m(\theta,\phi)
    \end{equation*}
    \begin{table}[h!]
        \centering
        \small
        \renewcommand{\arraystretch}{2.5}
        \begin{tabular}{llll}
            \toprule
            $n=1$ & $l=0$ & $m=0$ & $\psi_{100}=\dfrac{1}{\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\e[-\sigma]$\\
            $n=2$ & $l=0$ & $m=0$ & $\psi_{200}=\dfrac{1}{\sqrt{32\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(2-\sigma)\e[-\sigma/2]$\\
                  & $l=1$ & $m=0$ & $\psi_{210}=\dfrac{1}{\sqrt{32\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma\e[-\sigma/2]\cos\theta$\\
                  & $l=1$ & $m=\pm 1$ & $\psi_{21\pm 1}=\dfrac{1}{\sqrt{64\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma\e[-\sigma/2]\sin\theta\e[\pm i\phi]$\\
            $n=3$ & $l=0$ & $m=0$ & $\psi_{300}=\dfrac{1}{81\sqrt{3\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(27-18\sigma+2\sigma^2)\e[-\sigma/3]$\\
                  & $l=1$ & $m=0$ & $\psi_{310}=\dfrac{\sqrt{2}}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(6\sigma-\sigma^2)\e[-\sigma/3]\cos\theta$\\
                  & $l=1$ & $m=\pm 1$ & $\psi_{31\pm 1}=\dfrac{1}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(6\sigma-\sigma^2)\e[-\sigma/3]\sin\theta\e[\pm i\phi]$\\
                  & $l=2$ & $m=0$ & $\psi_{320}=\dfrac{1}{81\sqrt{6\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3](3\cos^2\theta-1)$\\
                  & $l=2$ & $m=\pm 1$ & $\psi_{32\pm 1}=\dfrac{1}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin\theta\cos\theta\e[\pm i\phi]$\\
                  & $l=2$ & $m=\pm 2$ & $\psi_{32\pm 2}=\dfrac{1}{162\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin^2\theta\e[\pm 2i\phi]$\\
            \bottomrule
        \end{tabular}
        \caption{The first few complete hydrogenlike atomic wave functions.}
        \label{tab:hydrogenWavefunctions}
    \end{table}
    \begin{itemize}
        \item Note that in Table \ref{tab:hydrogenWavefunctions}, $Z$ is the atomic number of the nucleus and $\sigma=Zr/a_0$ where $a_0$ is the Bohr radius.
        \item The orthonormality condition is
        \begin{equation*}
            \int_0^\infty\dd{r}r^2\int_0^\pi\dd{\theta}\sin\theta\int_0^{2\pi}\dd{\phi}\psi_{n'l'm'}^*(r,\theta,\phi)\psi_{nlm}(r,\theta,\phi) = \delta_{nn'}\delta_{ll'}\delta_{mm'}
        \end{equation*}
    \end{itemize}
    \item \textbf{Principal quantum number}: The quantum number $n$.
    \begin{itemize}
        \item $n=1,2,\dots$.
        \item The energy of the hydrogen atom depends on $n$.
    \end{itemize}
    \item \textbf{Angular momentum quantum number}: The quantum number $l$.
    \begin{itemize}
        \item $l=0,1,\dots,n-1$.
        \item The angular momentum of the electron about the proton is determined by $l$ through $|L|=\hbar\sqrt{l(l+1)}$.
        \item When $l=0,1,2,3$, it is customary to denote the value of $l$ with the letters $s,p,d,f$, respectively.
        \begin{itemize}
            \item After 3, we continue alphabetically past $f$.
            \item The origin of $s,p,d,f$ is historic from the designation of the observed spectral lines of atomic sodium; $s,p,d,f$ stand for sharp, principal, diffuse, and fundamental, respectively.
        \end{itemize}
        \item A wave function with $n=1$ and $l=0$ is called a $1s$ wave function; a wave function with $n=2$ and $l=0$ is called a $2s$ wave function, and so on and so forth.
    \end{itemize}
    \item \textbf{Magnetic quantum number}: The quantum number $m$.
    \begin{itemize}
        \item $m=0,\pm 1,\pm 2,\dots,\pm l$.
        \item So named because "the energy of a hydrogen atom in a magnetic field depends on $m$" \parencite[209]{bib:McQuarrieSimon}.
    \end{itemize}
    \item \textbf{Zeeman effect}: In the absence of a magnetic field, energy levels are $2l+1$-fold degenerate; in the presence of a magnetic field, however, the energy levels split.
    \begin{itemize}
        \item In the magnetic field case, $E$ is a function of both $n$ and $m$.
    \end{itemize}
    \item Since the radial wave functions are normalized, we have that
    \begin{equation*}
        \int_0^\infty[R_{nl}(r)]^2r^2\dd{r} = 1
    \end{equation*}
    \begin{itemize}
        \item It follows that the probability that the $nl$ electron lies between $r$ and $r+\dd{r}$ is
        \begin{equation*}
            \Prob = [R_{nl}(r)]^2r^2\dd{r}
        \end{equation*}
        \item Alternatively, we can consider the full wave functions to learn that
        \begin{equation*}
            \Prob = r^2\dd{r}\int_0^\pi\dd{\theta}\sin\theta\int_0^{2\pi}\dd{\phi}\psi_{nlm}^*(r,\theta,\phi)\psi_{nlm}(r,\theta,\phi)
        \end{equation*}
        \item For any specific example, these two equations will be in agreement.
    \end{itemize}
    \item The number of radial nodes is equal to $n-l-1$.
    \item \textbf{Orbital}: An atomic wave function.
    \item The above radial probability functions can be used to calculate average values of $r$ via
    \begin{equation*}
        \prb{r}_{nl} = \int_0^\infty rR_{nl}^*(r)R_{nl}(r)\dd{r}
    \end{equation*}
    \begin{itemize}
        \item Using general properties of the associated Laguerre polynomials, we can show that
        \begin{equation*}
            \prb{r}_{ns} = \frac{3}{2}a_0n^2
        \end{equation*}
    \end{itemize}
    \item We can calculate the most probable value of $r$ for an $nl$ wave function by differentiating the probability function and setting it equal to zero and then maximizing over the possible values.
    \item $p$ orbitals.
    \begin{itemize}
        \item Considers the $p_z$ orbital corresponding to $Y_1^0$, which is real.
        \item Covers tangent sphere picture, probability density plot, and contour map representations of atomic orbitals.
        \item When $m\neq 0$, the angular wave functions $Y_1^{\pm 1}$ are complex.
        \item To make them real, note that $|Y_1^1|^2=|Y_1^{-1}|^2$, so the probability densities and hence energies are the same.
        \item It follows that any linear combination of these two wave functions is another energy eigenfunction with the same energy, so take
        \begin{align*}
            p_x &= \frac{1}{\sqrt{2}}(Y_1^1+Y_1^{-1})&
            p_y &= \frac{1}{\sqrt{2}i}(Y_1^1-Y_1^{-1})
        \end{align*}
    \end{itemize}
    \item $d$ orbitals.
    \begin{itemize}
        \item We can take analogous linear combination to find real solutions (see Table \ref{tab:hydrogenRealWavefunctions}).
    \end{itemize}
    \begin{table}[h!]
        \centering
        \small
        \renewcommand{\arraystretch}{2.5}
        \begin{tabular}{llll}
            \toprule
            $n=1$ & $l=0$ & $m=0$ & $\psi_{1s}=\dfrac{1}{\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\e[-\sigma]$\\
            $n=2$ & $l=0$ & $m=0$ & $\psi_{2s}=\dfrac{1}{4\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(2-\sigma)\e[-\sigma/2]$\\
                  & $l=1$ & $m=0$ & $\psi_{2p_z}=\dfrac{1}{4\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma\e[-\sigma/2]\cos\theta$\\
                  & $l=1$ & $m=\pm 1$ & $\psi_{2p_x}=\dfrac{1}{4\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma\e[-\sigma/2]\sin\theta\cos\phi$\\
                  &       &           & $\psi_{2p_y}=\dfrac{1}{4\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma\e[-\sigma/2]\sin\theta\sin\phi$\\
            $n=3$ & $l=0$ & $m=0$ & $\psi_{3s}=\dfrac{1}{81\sqrt{3\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(27-18\sigma+2\sigma^2)\e[-\sigma/3]$\\
                  & $l=1$ & $m=0$ & $\psi_{3p_z}=\dfrac{\sqrt{2}}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(6\sigma-\sigma^2)\e[-\sigma/3]\cos\theta$\\
                  & $l=1$ & $m=\pm 1$ & $\psi_{3p_x}=\dfrac{1}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(6\sigma-\sigma^2)\e[-\sigma/3]\sin\theta\cos\phi$\\
                  &       &           & $\psi_{3p_y}=\dfrac{1}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}(6\sigma-\sigma^2)\e[-\sigma/3]\sin\theta\sin\phi$\\
                  & $l=2$ & $m=0$ & $\psi_{3d_{z^2}}=\dfrac{1}{81\sqrt{6\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3](3\cos^2\theta-1)$\\
                  & $l=2$ & $m=\pm 1$ & $\psi_{3d_{xz}}=\dfrac{\sqrt{2}}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin\theta\cos\theta\cos\phi$\\
                  &       &           & $\psi_{3d_{yz}}=\dfrac{\sqrt{2}}{81\sqrt{\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin\theta\cos\theta\sin\phi$\\
                  & $l=2$ & $m=\pm 2$ & $\psi_{3d_{x^2-y^2}}=\dfrac{1}{81\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin^2\theta\cos 2\phi$\\
                  &       &           & $\psi_{3d_{xy}}=\dfrac{1}{81\sqrt{2\pi}}\left( \dfrac{Z}{a_0} \right)^{3/2}\sigma^2\e[-\sigma/3]\sin^2\theta\sin 2\phi$\\
            \bottomrule
        \end{tabular}
        \caption{The first few complete real hydrogenlike atomic wave functions.}
        \label{tab:hydrogenRealWavefunctions}
    \end{table}
    \item "There is no fundamental reason to choose these linear combinations of spherical harmonics over the spherical harmonic themselves, but most chemists use the five $d$ orbitals given\dots because the functions\dots are real and have directional properties consistent with molecular structures" \parencite[218]{bib:McQuarrieSimon}.
    \item The Schr\"{o}dinger equation of the helium atom is
    \begin{equation*}
        \begin{split}
            &\left( -\frac{\hbar^2}{2M}\nabla^2-\frac{\hbar^2}{2m_e}\nabla_1^2-\frac{\hbar^2}{2m_e}\nabla_2^2 \right)\psi(\mathbf{R},\mathbf{r}_1,\mathbf{r}_2)\\
            &\quad +\left( -\frac{2e^2}{4\pi\epsilon_0|\mathbf{R}-\mathbf{r}_1|}-\frac{2e^2}{4\pi\epsilon_0|\mathbf{R}-\mathbf{r}_2|}+\frac{e^2}{4\pi\epsilon_0|\mathbf{r}_1-\mathbf{r}_2|} \right)\psi(\mathbf{R},\mathbf{r}_1,\mathbf{r}_2) = E)\psi(\mathbf{R},\mathbf{r}_1,\mathbf{r}_2)
        \end{split}
    \end{equation*}
    \begin{itemize}
        \item $\mathbf{R}$ is the position of the helium nucleus.
        \item $\mathbf{r}_1,\mathbf{r}_2$ are the positions of the two electrons.
        \item $M$ is the mass of the nucleus.
        \item $m_e$ is the electronic mass.
        \item $\nabla^2$ is the Laplacian operator with respect to the position of the nucleus.
        \item $\nabla_1^2,\nabla_2^2$ are the Laplacian operators with respect to the positions of the electrons.
    \end{itemize}
    \item The above Schr\"{o}dinger equation describes a three body problem and thus is not solveable.
    \item By fixing the nucleus at the center of a spherical coordinate system (as is justifiable since $M>>m_e$), we can simplify the Schr\"{o}dinger equation to
    \begin{equation*}
        -\frac{\hbar^2}{2m_e}(\nabla_1^2+\nabla_2^2)\psi(\mathbf{r}_1,\mathbf{r}_2)-\frac{2e^2}{4\pi\epsilon_0}\left( \frac{1}{r_1}+\frac{1}{r_2} \right)\psi(\mathbf{r}_1,\mathbf{r}_2)+\frac{e^2}{4\pi\epsilon_0|\mathbf{r}_1-\mathbf{r}_2|}\psi(\mathbf{r}_1,\mathbf{r}_2) = E\psi(\mathbf{r}_1,\mathbf{r}_2)
    \end{equation*}
    \item \textbf{Interelectronic repulsion term}: The $e^2/4\pi\epsilon_0|\mathbf{r}_1-\mathbf{r}_2|$ term in the Schr\"{o}dinger equation of the helium atom.
    \begin{itemize}
        \item This term makes it so that even this equation is still not solveable.
    \end{itemize}
\end{itemize}


\subsection*{Problems}
\begin{enumerate}[label={\textbf{6-\arabic*.}},ref={6-\arabic*}]
    \stepcounter{enumi}
    \item \label{prb:6-2}\marginnote{10/29:}In terms of the variable $\theta$, Legendre's equation is
    \begin{equation*}
        \sin\theta\dv{\theta}\left( \sin\theta\dv{\Theta(\theta)}{\theta} \right)+(\beta\sin^2\theta-m^2)\Theta(\theta) = 0
    \end{equation*}
    Let $x=\cos\theta$ and $P(x)=\Theta(\theta)$. Show that
    \begin{equation*}
        (1-x^2)\dv[2]{P(x)}{x}-2x\dv{P(x)}{x}+\left[ \beta-\frac{m^2}{1-x^2} \right]P(x) = 0
    \end{equation*}
    \begin{proof}[Answer]
        If $\cos\theta=\frac{x}{1}$, then
        \begin{equation*}
            \sin\theta = \sqrt{1-x^2}
        \end{equation*}
        Additionally,
        \begin{equation*}
            \dv{\Theta}{\theta} = \dv{P}{\theta}
            = \dv{P}{x}\dv{x}{\theta}
            = -\sin\theta\dv{P}{x}
            = -\sqrt{1-x^2}\dv{P}{x}
        \end{equation*}
        Therefore, we have by substitution and simplification that
        \begin{align*}
            0 &= \sqrt{1-x^2}\dv{\theta}\left( \sqrt{1-x^2}\cdot -\sqrt{1-x^2}\dv{P}{x} \right)+(\beta^2(1-x^2)-m^2)P(x)\\
            &= \sqrt{1-x^2}\dv{x}\left( -(1-x^2)\dv{P}{x} \right)\dv{x}{\theta}+(\beta^2(1-x^2)-m^2)P(x)\\
            &= (1-x^2)\dv{x}\left( (1-x^2)\dv{P}{x} \right)+(\beta^2(1-x^2)-m^2)P(x)\\
            &= (1-x^2)\left[ (1-x^2)\dv[2]{P}{x}-2x\dv{P}{x} \right]+(\beta^2(1-x^2)-m^2)P(x)\\
            &= (1-x^2)\dv[2]{P}{x}-2x\dv{P}{x}+\left[ \beta^2-\frac{m^2}{1-x^2} \right]P(x)
        \end{align*}
    \end{proof}
\end{enumerate}




\end{document}